\documentclass{article}
\usepackage{graphicx}
\usepackage{float}
\usepackage[utf8]{inputenc}
\title{Scale-up or scale-out: Is there one solution that fits all problem sizes?}
\date{}
\author{Marcel Stolin \\ marcelstolin@gmail.com}

\begin{document}


\maketitle


\begin{abstract}
Bla bla bla scaling is super wichtig.
This essay describes the scale-up and scale-out approache to scale a computing environment while mentioning their limits. 
%Since the beginning of Big Data, batch processing was the most popular choice for processing large amounts of generated data. These existing processing technologies are not suitable to process the large amount of data we face today. Research works developed a variety of technologies that focus on stream processing. Stream processing technologies bring significant performance improvements and new opportunities to handle Big Data. In this paper, we discuss the differences of batch and stream processing and we explore existing batch and stream processing technologies. We also explain the new possibilities that stream processing make possible.
\end{abstract}


% ==============================
% ###############################
% ==============================
\section{Introduction} \label{s_intro}
% Why is scaling needed (Moores Law + More powerful resources)
With the increase of interest in Artificial Intelligence and the growing complexity of modern computing architectures, more powerful resources are needed to adapt to business needs. If an IT environment reaches the limit of its computational power and can`t serve incoming requests efficiently, it has reached the limit of its scalability \cite{Wilder2012CloudPatterns}. The limit of scalability can be extended by providing more resources. Although, increasing the resources also increases the cost of running QUELLE.
% Introduction to scaling (horizontal and vertical)
The two primary approaches to increase the capacity of a computing environment are vertical scaling-up and horizontal scaling-out \cite{Abbott2015ScalabilityArt}. Both approaches are not exclusive and a computing environment can be designed to scale vertically, horizontally or both \cite{Wilder2012CloudPatterns}.
% Whats the problem (Adding resources or deploying components)
Although both strategies can be used to increase the capacity of a computing environment, each strategy follows a different approach.


% What is this essay about
The contribution of this essay is as follows. First, the approach of each scaling strategy is introduced. Second, based on a use case, the limitations of both scaling approaches are described, while mentioning the best approach to solve the problem. Third, a summary of this essay is given with a logical conclusion.


% ==============================
% ###############################
% ==============================
%\section{State of the art}
%There are many research studies around scalability of IT environments and different use-cases for horizontal and vertical scaling. Rossi et al. \cite{Rossi2019HorAndVert} created a solution to control the horizontal and vertical elasticity of container-based applications using Reinforcement learning  WHATS THE RESULT. 

%The cloud computing architecture has contributed to the scalability of modern computing environments. Cloud computing makes it easy to increase the computing capacity of a computing environment on demand \cite{Mahmood2013CloudConcepts}.


% ==============================
% ###############################
% ==============================
\section{Background}

This section introduces the concept of the scaling-up and scaling-out approaches.

\subsection{Scale-up}
% Explain
Scale-up refers to improving the hardware of individual nodes \cite{Wilder2012CloudPatterns}. By adding more powerful resources to a node, a node can take more throughput and perform more specialized tasks \cite{Abbott2015ScalabilityArt}.
% How can it be achieved
An increase of the computing capacity of a node can include adding memory or adding more CPU cores \cite{Wilder2012CloudPatterns}.
Due to the low complexity of scaling-up, it is the most applied scaling approach \cite{Wilder2012CloudPatterns}.


\subsection{Scale-out}
% Explain
Scaling-out is achieved by adding more nodes to the computing environment \cite{Wilder2012CloudPatterns}. Adding more nodes, increases the overall computing capacity of the environment and in addition, the workload can be distributed across all nodes to handle and increasing number of requests \cite{Wilder2012CloudPatterns, Abbott2015ScalabilityArt}.
% Homegenereous
To scale efficiently, nodes should be homogeneous. Homogeneous nodes are able to perform the same work and will respond as other nodes \cite{Abbott2015ScalabilityArt}.
% How can it be achieved
%In cloud computing, instances are typically deployed using container technologies. Cloning containers is easy due to their lightweight architecture \cite{Abbott2015ScalabilityArt}.


% Example
A typical example for using a scale-out strategy is, to create new instances of a database to serve read-operations for rapidly increasing requests.


% ==============================
% ###############################
% ==============================
\section{Scale-up or scale-out}
\paragraph{}
This section describes a typical use case where scaling is necessary. It gives an overview about the limitations of each scaling approach and concludes with the most logical scaling approach.


\paragraph{}
% The case
If a computing environment reaches the limitations of its hardware capacity, options are to scale-up by adding more powerful resources or to scale-out to distribute the work across multiple nodes. If a computing environment is not designed to scale-out, the only option remaining is to add more powerful hardware \cite{Abbot2011ScalabilityRules}.
% Problem with scaling-up
Scaling-up by adding more powerful hardware to the computing environment, is limited by maximum hardware capacity and causes downtimes due to IT resource replacements \cite{Mahmood2013CloudConcepts}. An economical threshold can be reached, where buying more powerful hardware is not affordable. Another limitation can be that no more powerful hardware is available by the provider \cite{Abbot2011ScalabilityRules, Wilder2012CloudPatterns}.
% Problems solved with scaling-out?
If the environment is designed to scale-out, buying more powerful resources is not needed. The computing capacity can be increased by adding more nodes which are instantly available. In this case, the scaling-out strategy gives an IT environment the flexibility to scale without increasing the cost \cite{Abbot2011ScalabilityRules}.


% About scaling-out
Designing an environment to scale-out, shifts the focus from infrastructure to development. The goal of scaling-out is to increase the computing capacity by combining the computing power of several nodes, which makes a scale-out design more complex compared to a scale-up architecture \cite{Wilder2012CloudPatterns}. The computing capacity of a scale-out architecture is limited by the computing power of each additional node. Therefore, scaling-out is more efficient with homogeneous nodes, where each node adds the same amount of computing power to the system \cite{Wilder2012CloudPatterns}. With non-homogeneous nodes, the complexity of capacity planning and distributing work across nodes grows.


% Summary
Scaling-up should be considered than scaling-out due to its significant drawbacks. In general, computing environments should be designed to scale-out in the first place. While the cost of buying more powerful hardware grows proportionally, the cost of investing engineering power to design an environment to scale-out grows linear \cite{Abbot2011ScalabilityRules}. With a scaling-out architecture, the environment can adapt to business needs without interruptions. If the limitation of a scaling-out architecture is reached, scaling-up is still an remaining option.


% ==============================
% ###############################
% ==============================
\section{Conclusion}
The ability to scale is a critical characteristic of modern and complex computing environments. With the ability to scale, a computing environment is able to adapt to business needs. Scaling-up and scaling-out are two scaling approaches to increase the computing capacity of a computing environment. Scaling-up is the simpler approach, but it is limited and comes with significant disadvantages according to cost and flexibility. On the other hand, scaling-out adds more complexity to the environment but is able to scale the environment without increasing the cost proportionally. A modern IT environment should be designed to scale-out in the first place to handle performance spikes and keep the cost low.


% ==============================
% ###############################
% ==============================
\pagebreak
\bibliographystyle{plain}
\bibliography{mybib}

\end{document}